Parsing Postfix log files

### 2007-06-26 doc/logparser.tex
Improved the graphs and explanation for the dips in normal/shuffle/reverse; added a better graph and removed the two it replaced. Added all the files necessary for rebuilding - generating some of them again would be painful, and the explanations are tied to the log files used now. Removed unnecessary files.

XXX LOTS MORE REFERENCES

XXX MAYBE ADD A SECTION ABOUT OPTIMISATION OF THE DATABASE\@.

@@

The parser does not try to detect overlapping rules;\footnote{It may be
possible to parse each rule's regex, generate an NFA, and compare all NFAs
to determine if any overlap.  The author has not attempted to do this: such
a project would probably qualify for a PhD, and may involve solving the
Halting Problem and circumventing the Church-Turing Thesis.} that
responsibility is left to the author of the rules.  Unintentionally
overlapping rules lead to inconsistent parsing and data extraction because
the order in which rules are tried against each line determines which rule
matches.  Overlapping rules are frequently a requirement, allowing a more
specific rule to match some lines and a more general rule to match the
remainder, e.g. separating smtp delivery to a filtering proxy from smtp
delivery to the rest of the world.  The algorithm provides a facility for
handling overlapping rules: the priority field in each rule (defaults to
zero).  Rules are tried in order of priority, highest first, overriding the
rule\_order field maintained by the parser.  Negative priorities may also
be useful for catchall rules.

@@

    \item Sort the regexs and visually inspect the list, e.g. with SQL
        similar to: \textbf{select regex from rules order by regex;}

@@

XXX WRITE ABOUT DATABASE OPTIMISATION??

@@

XXX THIS NEXT PARAGRAPH MIGHT NEED TO BE REWRITTEN\@.

The dips at logfiles 22 and 62--68 correspond to peaks in log file size in
graph~\ref{execution time vs file size vs number lines}.  The explanation
took some time to arrive at, but turns out to be reasonably simple.
Examining the database after parsing logfile 22 shows 64 rules matched
during parsing, but the distribution of matching rules is extremely skewed,
as shown in graph~\ref{log22 distribution by rule}.  Graph~\ref{rule
distribution by program} shows rule distribution by program.  Comparing the
number of log lines per program and rules per program shows that
\textit{postfix/qmgr\/} has the highest number of log lines but has a low
number of rules to try, so the average execution time for a
\textit{postfix/qmgr\/} is quite low, reducing the overall average
execution time per log line.

XXX CREATE A GRAPH AVERAGE EXECUTION TIME PER LOG FILE\@.

@@

\subsubsection{Log 22 distribution by program}
\label{log22 distribution by program}
\includegraphics{plot-log22-distribution-by-program.ps}

\subsubsection{Rule distribution by program}
\label{rule distribution by program}
\includegraphics{plot-log22-rule-count-by-program.ps}

>>>>

@@

The parser does not try to detect overlapping rules;\footnote{It may be
possible to parse each rule's regex, generate an NFA, and compare all NFAs
to determine if any overlap.  The author has not attempted to do this: such
a project would probably qualify for a PhD, and may involve solving the
Halting Problem and circumventing the Church-Turing Thesis.} that
responsibility is left to the author of the rules.  Unintentionally
overlapping rules lead to inconsistent parsing and data extraction because
the order in which rules are tried against each line determines which rule
matches.  Overlapping rules are frequently a requirement, allowing a more
specific rule to match some lines and a more general rule to match the
remainder, e.g.\ separating smtp delivery to a filtering proxy from smtp
delivery to the rest of the world.  The algorithm provides a facility for
handling overlapping rules: the priority field in each rule (defaults to
zero).  Rules are tried in order of priority, highest first, overriding the
rule\_order field maintained by the parser.  Negative priorities may also
be useful for catchall rules.

@@

    \item Sort the regexs and visually inspect the list, e.g.\ with SQL
        similar to: \textbf{select regex from rules order by regex;}

@@

@@

The dips at logfiles 22 and 62--68 correspond to peaks in log file size in
graph~\ref{execution time vs file size vs number lines}.  The explanation
for this took some time to arrive at, but it turns out to be reasonably
simple.  Examining the database after parsing logfile~22 shows 64 rules
matched during parsing, but the distribution of matching rules is extremely
skewed, as shown in graph~\ref{log22 distribution by rule}.  Comparing the
number of log lines per program versus rules per program in
graph~\ref{log22 line and rule distribution by program} shows that
\textit{postfix/qmgr\/} has the highest number of log lines; however it has
a low number of rules, so the average parsing time for a
\textit{postfix/qmgr\/} log line is quite low, reducing the overall average
parsing time per log line.

XXX FIGURE OUT THE DIP AT LOG 34, mail.log-20070228

@@

\subsubsection{Log 22 line and rule distribution by program}
\label{log22 line and rule distribution by program}
\includegraphics{plot-log22-distribution-versus-rule-count.ps}
